Assessing the impacts of time to detection distribution assumptions on detection probability estimation

### 2017-02-21 ReviewerComments.tex
Add comments

\newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
\newcommand{\jarad}[1]{{\color{Orange} JARAD: \#1}}

@@

I enjoyed reading this well-written manuscript, and believe that the modeling framework proposed is sufficiently versatile to be used in a wide range of applications. In addition to the points made by the referees, I have two comments:
\begin{enumerate}
\item Reviewer 1 is not convinced that it is helpful to make explicit the link to survival analysis. I tend to disagree with this, since I think that this is what renders the approach widely applicable and also extendable. In fact, the general idea strongly reminded me of several papers published recently by David Borchers and colleagues (using concepts from survival analysis to model detection events within distance sampling and spatially explicit capture-recapture). I was wondering if there is a common denominator between his work and the present manuscript (but the authors need not necessarily explore this).
\begin{quote}
We agree, this manuscript has strong parallels to the continuous-time models of David Borchers and colleagues.  
Their models are for removal-distance data, whereas ours is for removal-only data.
\end{quote}
\item In Section 3.4, models are built where the various TTDDs considered are linked to covariates via the rate parameter. It was not clear to me why the rate parameter was chosen to be modeled as a function of covariates, as opposed to the mean of the TTDD. For example, in gamma GLMs, the mean is modeled as a function of covariates, with the shape parameter fixed (the latter as done also in the present work). Basically I'm now wondering a) why the rate rather than the mean was modeled, and b) if that matters at all. In 3.1, a justification seems to be given, but I do not follow the reasoning here – why would the inclusion of random and fixed effects be any different if not the rate but the mean would be used? Sorry if I'm just missing the obvious.
\begin{quote}
We have added a statement clarifying that there is no mathematical difference.
\adam{
In response to (a), while we fit our models with common distributions for $f_T(t)$, we could alternatively begin with a detection hazard rate $h(t)$.
Given this context, the detection rate feels more like the focus of inference than the expected time to detection.
}
\end{quote}
\end{enumerate}

@@

Line 14: Perhaps instead of saying its not justified, mention it can lead to bias
\begin{quote}
Done.
\end{quote}

@@

3.1 \\
Line 60: In the context of estimating abundance of wildlife, justifying the inclusion of the Weibull and lognormal distributions because they are often used in survival analysis is weak.  Are there biologically plausible scenarios in which these distributions are appropriate given this type of data?  Justify here similar to the Gamma distribution.  Otherwise, omit.
\begin{quote}
We have added an example where detection rates vary across individuals, causing the marginal detection rate to decline over time.
\end{quote}

Simulation Analyses\\
Line 57-59: This is the first mention of peaked and non-peaked TTDDs.  Please define these terms and how they are manifested during point-count surveys in the Introduction - including a biological justification that warrants their inclusion in simulation studies.
\begin{quote}
We have made the meaning of `peaked' and `nonpeaked' more explicit in Section \ref{M-sec:sim}.
We do not believe this distinction requires a separate explanation in the Introduction.
\end{quote}

As a potential user of such a model presented in this manuscript, I find the simulation analyses somewhat lacking.   
\begin{enumerate}
\item Users will most certainly not bother using such complex modeling approach to model an intercept only - and thus, evaluating TTDDs based on intercept-only models seems overly simple and not very informative.
\begin{quote}
We agree that an intercept-only model is simplistic and will not be of much practical use in applied settings.
However, it can be informative.
TTDDs that are biased (e.g. non-mixture lognormal) or not robust to misspecification (e.g. mixture exponential) may not be trustworthy in more complicated analyses.
\end{quote}
\item Simulations incorporating variable sample sizes to evaluate relative bias and precision would be very useful in terms of study design and applicability.  
\begin{quote}
\adam{Investigate more/different simulations}
\end{quote}
\item How might study design impact bias and precision?  For example, length of survey relative to peaked TTDDs and pooling of time periods during data collection?  These are also scenarios that would benefit readers.  Alternatively, these issues could be addressed in the Discussion.
\begin{quote}
\adam{We have not investigated either of these issues.
During my Masters work, we examined the effects of censoring on a mixture exponential model and found a 33\% increase in credible interval widths for $\pdet$ when switching from 9- to 3-interval data in a model with covariates.
However, I would not dream of extending those findings to other distributions.
We obviously would prefer longer durations, but those risk the violation of closure assumptions.
\citet{Petit1995, Johnson2008, LeeMarsden2008} and \citet{Reidy2011} all address the effects of duration.
I am inclined not to add Discussion on topics about which I do not have well-informed opinions.
}
We have added a mention of the trade-offs of a settling down period to reduce observer effects on behavior.
\end{quote}

\item Providing computation times for scenarios that vary by model complexity and sample size (including the computer used for analyses) may also provide a gauge for users as to whether this method is feasible for their needs.
\begin{quote}
\adam{Because of improvements in Stan, I want to run more simulation reps before quantifying this.}
\end{quote}
\item I don’t fully appreciate the need for both 50\% and 90\% coverage estimates.  It seems one estimate would be sufficient, and typically 90\% or 95\% credible intervals are used with these types of analyses?
\begin{quote}
We have retained just the 50\% coverage.
\adam{Waiting until all new simulations are run.}
\end{quote}
\end{enumerate}

Results\\
Page 18, Figure 2:  Holy moly, the CIs of the mixture distributions are exceedingly wide given a sample size of 381 detections (which is a lot relative to many studies).
\begin{quote} Clarification: 947 detections at 381 sites (which is \textit{really} a lot of detections)\end{quote}
For example, The Gamma distribution, touted as the most accurate based on the simulation study, suggests 95\% CIs of p ranges from ~ 0.3 to 1?  And subsequently CIs around N are from ~1.75 to 3.25 so we can say there are somewhere between 56 and 1,778  birds?
Is that a useful estimate?  There is a tradeoff between accuracy and precision happening here, and although precise, biased estimates are not what we want, any alternative needs to provide precision at levels that are still informative.  This is a limitation not sufficiently addressed in the Discussion.
\begin{quote} 
We think our figure may have been unclear.
The cited credible intervals are for \textit{uncounted} birds not total abundance.
We have now changed the figure to $\log_{10}$(Abundance) to avoid confusion.
The gamma TTDD 95\% credible interval for total abundance is (1007, 2836).
This two- to three-fold scale of uncertainty is commensurate with other studies cited in this manuscript \citep{Diefenbach2007, Reidy2011, Solymos2013, Amundson2014}.
\end{quote}

Discussion\\
Page 21, Lines 12-16:  I do not agree with this assessment.  I think these models are appropriate for a particular subset of data - single species discrete-count surveys with a large sample size and relatively low availability (i.e., whether the animal provides a cue to an observer during the survey)   
\begin{quote}
\adam{What is a `discrete-count survey'?  I have not found a clear reference.\\
I accede the `large sample' point.  But based on our findings, I'd advocate against modeling abundance/perception from any small dataset, because analysis requires either a strong assumption about detection rates or a strong prior.  Any comment on this point would be strengthened if we ran small-dataset simulations.\\
I think the low-availability stipulation is redundant.  If availability is really high, then a removal analysis is not necessary at all.}
\end{quote}
where non-constant detection through time is suspected based on the behavioral ecology of the species.
\begin{quote}
\adam{What I do not advocate is defaulting to the constant-rate assumption: by the precautionary principle, the burden of proof should rest with the analyst who wishes to assume constant detection rather than with the analyst who wishes to account for non-constant detection.}
\end{quote}

As the authors’ suggest, most point-count studies are designed to maximize availability, which is typically high for birds (e.g., 80-95\%) with perceptibility (e.g., the probability an observer detects a cue that is given by an animal) being the much larger source of bias.  Therefore, this model, without including a supplemental method to estimate perceptibility, is of limited use to many readers.
\begin{quote}
We agree.  Our current research addresses this issue in more depth.
\end{quote}
Further, although the method may be more accurate even for constant detection data, the loss in precision is such that users will likely not opt to use it unless they suspect non-constant detection through time.  Further, given its presumably arduous computation times (albeit unknown), adding a perceptibility component to the model would likely greatly increase computational demands with unknown implications to bias and precision.  Including a paragraph in the Discussion that outlines when the model would be most beneficial to users (e.g., what type of data, surveys, or sample sizes) and explicitly stating drawbacks (e.g., computation time, loss of precision) would greatly improve this section.  
\begin{quote}
\adam{This last recommendation seems feasible, though I think our final paragraph already addesses some of these issues.}
\end{quote}
\vspace{1.5cm}

@@

Diagnostics: Coverage. The idea of testing coverage is fantastic, but poorly implemented. First, the number of simulations (16 in Table 1) is exceedingly low to be able to reliably diagnose the patterns in coverage or the statistical explanation of these patterns. Second, the statistical properties of the estimator of p(det) are likely depending on the size of the true value of p(det). Therefore, I think Table 1 should have been repeated for a whole range of values of the true value of p(det)
\begin{quote}
\adam{We should run more simulations.  Reviewer 1 wants more variation in $n^{(obs)}$.  Reviewer 2 wants a wide range or simulated $\pdet$.}
\end{quote}
(Is a boundary like the one in Olkin et al 1981, for your case, a hard one such that below it estimation is bad, and above it estimation is ``uniformly" good?).
\begin{quote}
It is not a hard boundary.  It results from the relative sample mean and sample variance of large samples.
However, similar thresholds continue to surface again and again in various abundance-estimating contexts: 
\begin{itemize}
\item \citet{Veech2016}: suggest  $p>0.5$ in the context of individual-level detection effects (no time-to-detection data). 
\item \citet{Field2016}: plots indicate uncertainty in $\hat{N}$ increase rapidly as $\pdet$ decreases below 0.40.  Methods include removal, double-observer, distance, and multiple-visit.
\item \citet{Davis2016}: report accurate $\hat{N}$ at $\pdet >0.40$ in a traditional removal-sampling context, though for small abundance ($N<50$) they required $\pdet > 0.7$.
\end{itemize}
\adam{Jarad, I am unclear on the purpose of the Olkin-citing paragraph.  I don't feel like the above arguments really merit inclusion in the text.}
\end{quote}

@@

Diagnostics: relation of the target parameter to other parameters in the models. Precisely because, as the authors mention, the statistical properties of the estimator of p(det) are tightly linked with the statistical properties of the abundance estimator, then the author(s) should detail explicitly how a bad coverage in p(det), for instance, affects the estimation of abundances. The author(s) mention that bias in p translates into a bias in abundance estimation, I would like to see diagnostics for the realized N values too. And to do that, two words come to mind: profile likelihoods. Unless you are explicitly adopting a subjective Bayesian approach, you basically declare a priori ignorance for your parameters.  Furthermore, you state that typically you will be in a case where data sets aren't large, so the information in the data is not ``swamping" the priors, so to speak.
\begin{quote}
We have posterior estimates of abundance from our model fits.  We have added them to the manuscript.\\
\adam{Thoughts on how to do this:
\begin{enumerate}
\item Generate Tables 1 \& 2 for $\hat{N}$ as well as $\pdet$, but this gets crowded.
\item Supplement Tables 1 \& 2 with some manner of confidence intervals for $\hat{N} / N$, but averaging across replicates is not a trivial issue
\item Supplement Figure 1 with a row of representative abundance posteriors (no averaging required)
\item Following the profile likelihood comment, generate a plot of $\hat{N}$ vs. $\pdet$ for a representative sample.
\end{enumerate}
}
\end{quote}

Lele et al 2010 (a frequent co-author of Solymos) propose a pretty neat diagnostic tool to assess
estimability that is the by product of tricking a bayesian MCMC set up into Maximum Likelihood estimation for hierarchical models (see Lele et al 2007, 2010 and others. The keyword is ``Data Cloning"). You are one step away from using Solymos and Lele's ``Data Cloning" (DC) approach to get the full ML estimation working (See Lele et al 2010 and other Data Cloning papers). Now, this is relevant because using DC you can provide clear estimability diagnostics for the parameters of interest.  Are there parameters that are technically not estimable?  Given that you don't have informative priors, you are ``driving in the dark" so to speak if you are not certain if some of your parameters are un-identifiable (see Lele et al 2010 and discussion in Lele and Dennis 2009).  Note that I am NOT asking you to re-do all the analysis using a ML approach, I am just suggesting the fact that using DC to get the ML estimates one can, as a very useful by product, get very neat diagnostics regarding
the estimability of your parameters.  Again, see Lele et al 2010.  Implementing DC would only imply a simple modification to the programs you already have. And by the way, Solymos has a DC package easy to use.
\begin{quote}
\adam{
\begin{itemize}
\item The Solymos package is built for BUGS and JAGS.  Most of the tools are easily hand-coded, I think.
\item To simplify our lives, we could choose to attempt this for only quick-running models on the actual data... there's no reason to suspect identifiability issues for the gamma TTDD that would not appear for the Weibull or lognormal.
\item Technical note: they suggest using narrower priors as the number of clones increases.  Likewise, we could input posterior means as initial values.
\end{itemize}
}
\end{quote}

@@

One thing that the authors could do is to take the reader by the hand by building equation 1 bit by bit: present first the simplest version of equation 1 and what is customary to date, and then add modifications/hierarchies plus text and build the model with explicit equations and little by little arrive to the final product, which is the current equation 1. Nothing better than clear simple math aided by short explanations to present a model unambiguously.
\begin{quote}
\adam{
This potentially requires a lot of rewiring.  Two options are: (i) highlight each component of Equation (1) as we address it in the text, because we do address them all, or (ii) at the end of Section 3.2, provide a simpler version of the model without mixtures and without covariates.
}
\end{quote}

@@

The authors could have crafted a beautiful and very telling graph (Figure 1) representing the multinomial intervals, the TTDD, the right censoring due to the end of the observation period, and the increased detection probability mixture component. Such figure (which is what I did by hand as I was reading the manuscript) would be a very useful guide in the reading of the paper.
\begin{quote}
We have crafted a beautiful and very telling graph (Figure 1).
\end{quote}

@@

@@

>>>>

\newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
\newcommand{\jarad}[1]{{\color{Orange} JARAD: #1}}

@@

I enjoyed reading this well-written manuscript, and believe that the modeling framework proposed is sufficiently versatile to be used in a wide range of applications. In addition to the points made by the referees, I have two comments:
\begin{enumerate}
\item Reviewer 1 is not convinced that it is helpful to make explicit the link to survival analysis. I tend to disagree with this, since I think that this is what renders the approach widely applicable and also extendable. In fact, the general idea strongly reminded me of several papers published recently by David Borchers and colleagues (using concepts from survival analysis to model detection events within distance sampling and spatially explicit capture-recapture). I was wondering if there is a common denominator between his work and the present manuscript (but the authors need not necessarily explore this).
\begin{quote}
We agree, this manuscript has strong parallels to the continuous-time models of David Borchers and colleagues.  
Their models are for removal-distance data, whereas ours is for removal-only data.
\jarad{They also don't look at other distributions for the removal portion.}
\end{quote}

\item In Section 3.4, models are built where the various TTDDs considered are linked to covariates via the rate parameter. It was not clear to me why the rate parameter was chosen to be modeled as a function of covariates, as opposed to the mean of the TTDD. For example, in gamma GLMs, the mean is modeled as a function of covariates, with the shape parameter fixed (the latter as done also in the present work). Basically I'm now wondering a) why the rate rather than the mean was modeled, and b) if that matters at all. In 3.1, a justification seems to be given, but I do not follow the reasoning here – why would the inclusion of random and fixed effects be any different if not the rate but the mean would be used? Sorry if I'm just missing the obvious.
\begin{quote}
We have added a statement clarifying that there is no mathematical difference.
\adam{
In response to (a), while we fit our models with common distributions for $f_T(t)$, we could alternatively begin with a detection hazard rate $h(t)$.
Given this context, the detection rate feels more like the focus of inference than the expected time to detection.
}
\end{quote}
\end{enumerate}

@@

Line 14: Perhaps instead of saying its not justified, mention it can lead to bias
\begin{quote}
Done. \jarad{Be more explicit, e.g. "We've added the following to Section X.x" ...}
\end{quote}

@@

3.1 \\
Line 60: In the context of estimating abundance of wildlife, justifying the inclusion of the Weibull and lognormal distributions because they are often used in survival analysis is weak.  Are there biologically plausible scenarios in which these distributions are appropriate given this type of data?  Justify here similar to the Gamma distribution.  Otherwise, omit.
\begin{quote}
We have added an example \jarad{where?} where detection rates vary across individuals, causing the marginal detection rate to decline over time.
\end{quote}

Simulation Analyses\\
Line 57-59: This is the first mention of peaked and non-peaked TTDDs.  Please define these terms and how they are manifested during point-count surveys in the Introduction - including a biological justification that warrants their inclusion in simulation studies.
\begin{quote}
We have made the meaning of `peaked' and `nonpeaked' more explicit in Section \ref{M-sec:sim}.
We do not believe this distinction requires a separate explanation in the Introduction.
\jarad{Can we reword to remove peaked and nonpeaked here?}
\end{quote}

As a potential user of such a model presented in this manuscript, I find the simulation analyses somewhat lacking.   
\begin{enumerate}
\item Users will most certainly not bother using such complex modeling approach to model an intercept only - and thus, evaluating TTDDs based on intercept-only models seems overly simple and not very informative. \jarad{Computation time.}
\begin{quote}
We agree that an intercept-only model is simplistic and will not be of much practical use in applied settings.
However, it can be informative.
TTDDs that are biased (e.g. non-mixture lognormal) or not robust to misspecification (e.g. mixture exponential) may not be trustworthy in more complicated analyses.
\end{quote}
\item Simulations incorporating variable sample sizes to evaluate relative bias and precision would be very useful in terms of study design and applicability.  
\begin{quote}
\adam{Investigate more/different simulations}
\end{quote}
\item How might study design impact bias and precision?  For example, length of survey relative to peaked TTDDs and pooling of time periods during data collection?  These are also scenarios that would benefit readers.  Alternatively, these issues could be addressed in the Discussion.
\begin{quote}
\adam{We have not investigated either of these issues.
During my Masters work, we examined the effects of censoring on a mixture exponential model and found a 33\% increase in credible interval widths for $\pdet$ when switching from 9- to 3-interval data in a model with covariates.
However, I would not dream of extending those findings to other distributions.
We obviously would prefer longer durations, but those risk the violation of closure assumptions.
\citet{Petit1995, Johnson2008, LeeMarsden2008} and \citet{Reidy2011} all address the effects of duration.
I am inclined not to add Discussion on topics about which I do not have well-informed opinions.
}
We have added a mention of the trade-offs of a settling down period to reduce observer effects on behavior.
\end{quote}

\item Providing computation times for scenarios that vary by model complexity and sample size (including the computer used for analyses) may also provide a gauge for users as to whether this method is feasible for their needs.
\begin{quote}
\adam{Because of improvements in Stan, I want to run more simulation reps before quantifying this.}
\end{quote}

\item I don’t fully appreciate the need for both 50\% and 90\% coverage estimates.  It seems one estimate would be sufficient, and typically 90\% or 95\% credible intervals are used with these types of analyses?
\begin{quote}
We have retained just the 50\% coverage.
\adam{Waiting until all new simulations are run.}
\jarad{I think we can explain why here, but because the other reviewer also 
mentioned ``poor implementation'' of coverage, we may need to address why in the manuscript.
I think computation time is a reason.}
\end{quote}
\end{enumerate}

Results\\
Page 18, Figure 2:  Holy moly, the CIs of the mixture distributions are exceedingly wide given a sample size of 381 detections (which is a lot relative to many studies).
\begin{quote} Clarification: 947 detections at 381 sites (which is \textit{really} a lot of detections)\end{quote}
For example, The Gamma distribution, touted as the most accurate based on the simulation study, suggests 95\% CIs of p ranges from ~ 0.3 to 1?  And subsequently CIs around N are from ~1.75 to 3.25 so we can say there are somewhere between 56 and 1,778  birds?
Is that a useful estimate?  There is a tradeoff between accuracy and precision happening here, and although precise, biased estimates are not what we want, any alternative needs to provide precision at levels that are still informative.  This is a limitation not sufficiently addressed in the Discussion.
\begin{quote} 
We think our figure may have been unclear.
The cited credible intervals are for \textit{uncounted} birds not total abundance.
We have now changed the figure to $\log_{10}$(Abundance) to avoid confusion.
The gamma TTDD 95\% credible interval for total abundance is (1007, 2836).
This two- to three-fold scale of uncertainty is commensurate with other studies cited in this manuscript \citep{Diefenbach2007, Reidy2011, Solymos2013, Amundson2014}.
\jarad{This is the same order of magnitude and therefore is useful in my opinion.}
\end{quote}

Discussion\\
Page 21, Lines 12-16:  I do not agree with this assessment.  I think these models are appropriate for a particular subset of data - single species discrete-count surveys with a large sample size and relatively low availability (i.e., whether the animal provides a cue to an observer during the survey)   
\begin{quote}
\adam{What is a `discrete-count survey'?  I have not found a clear reference.\\
I accede the `large sample' point.  But based on our findings, I'd advocate against modeling abundance/perception from any small dataset, because analysis requires either a strong assumption about detection rates or a strong prior.  Any comment on this point would be strengthened if we ran small-dataset simulations.\\
I think the low-availability stipulation is redundant.  If availability is really high, then a removal analysis is not necessary at all.}
\jarad{But how do you know availability is high?}
\end{quote}
where non-constant detection through time is suspected based on the behavioral ecology of the species.
\begin{quote}
\adam{What I do not advocate is defaulting to the constant-rate assumption: by the precautionary principle, the burden of proof should rest with the analyst who wishes to assume constant detection rather than with the analyst who wishes to account for non-constant detection.}
\jarad{Agreed.}
\end{quote}

As the authors’ suggest, most point-count studies are designed to maximize availability, which is typically high for birds (e.g., 80-95\%) with perceptibility (e.g., the probability an observer detects a cue that is given by an animal) being the much larger source of bias.  Therefore, this model, without including a supplemental method to estimate perceptibility, is of limited use to many readers.
\begin{quote}
We agree.  Our current research addresses this issue in more depth.
\end{quote}
Further, although the method may be more accurate even for constant detection data, the loss in precision is such that users will likely not opt to use it unless they suspect non-constant detection through time.  Further, given its presumably arduous computation times (albeit unknown), adding a perceptibility component to the model would likely greatly increase computational demands with unknown implications to bias and precision.  Including a paragraph in the Discussion that outlines when the model would be most beneficial to users (e.g., what type of data, surveys, or sample sizes) and explicitly stating drawbacks (e.g., computation time, loss of precision) would greatly improve this section.  
\begin{quote}
\adam{This last recommendation seems feasible, though I think our final paragraph already addesses some of these issues.}
\jarad{If we already address it, then you should just quote from the manuscript.}
\end{quote}
\vspace{1.5cm}

@@

Diagnostics: Coverage. The idea of testing coverage is fantastic, but poorly implemented. First, the number of simulations (16 in Table 1) is exceedingly low to be able to reliably diagnose the patterns in coverage or the statistical explanation of these patterns. Second, the statistical properties of the estimator of p(det) are likely depending on the size of the true value of p(det). Therefore, I think Table 1 should have been repeated for a whole range of values of the true value of p(det)
\begin{quote}
\adam{We should run more simulations.  Reviewer 1 wants more variation in $n^{(obs)}$.  Reviewer 2 wants a wider range of simulated $\pdet$.}
\end{quote}
(Is a boundary like the one in Olkin et al 1981, for your case, a hard one such that below it estimation is bad, and above it estimation is ``uniformly" good?).
\begin{quote}
It is not a hard boundary.  It results from the relative sample mean and sample variance of large samples.
However, similar thresholds continue to surface again and again in various abundance-estimating contexts: 
\begin{itemize}
\item \citet{Veech2016}: suggest  $p>0.5$ in the context of individual-level detection effects (no time-to-detection data). 
\item \citet{Field2016}: plots indicate uncertainty in $\hat{N}$ increase rapidly as $\pdet$ decreases below 0.40.  Methods include removal, double-observer, distance, and multiple-visit.
\item \citet{Davis2016}: report accurate $\hat{N}$ at $\pdet >0.40$ in a traditional removal-sampling context, though for small abundance ($N<50$) they required $\pdet > 0.7$.
\end{itemize}
\adam{Jarad, I am unclear on the purpose of the Olkin-citing paragraph.  I don't feel like the above arguments really merit inclusion in the text.}
\jarad{In my opinion, if $\pdet$ is low, then information about $\pdet$ and abundance are 
obtained through the priors and not through the data. These citations as well as the Olkin paragraph
indicate that statistical methods cannot help us to recover when $\pdet$ is low unless we 
are willing to make strong assumptions.}
\end{quote}

@@

\jarad{Although we should look at more simulations. The purpose of the current values
was to give some credence to the Ovenbird analysis.}

Diagnostics: relation of the target parameter to other parameters in the models. Precisely because, as the authors mention, the statistical properties of the estimator of p(det) are tightly linked with the statistical properties of the abundance estimator, then the author(s) should detail explicitly how a bad coverage in p(det), for instance, affects the estimation of abundances. The author(s) mention that bias in p translates into a bias in abundance estimation, I would like to see diagnostics for the realized N values too. And to do that, two words come to mind: profile likelihoods. Unless you are explicitly adopting a subjective Bayesian approach, you basically declare a priori ignorance for your parameters.  Furthermore, you state that typically you will be in a case where data sets aren't large, so the information in the data is not ``swamping" the priors, so to speak.
\begin{quote}
We have posterior estimates of abundance from our model fits.  We have added them to the manuscript.\\
\adam{Thoughts on how to do this:
\begin{enumerate}
\item Generate Tables 1 \& 2 for $\hat{N}$ as well as $\pdet$, but this gets crowded.
\item Supplement Tables 1 \& 2 with some manner of confidence intervals for $\hat{N} / N$, but averaging across replicates is not a trivial issue
\item Supplement Figure 1 with a row of representative abundance posteriors (no averaging required)
\item Following the profile likelihood comment, generate a plot of $\hat{N}$ vs. $\pdet$ for a representative sample.
\end{enumerate}
}
\jarad{I think 3.}
\end{quote}

Lele et al 2010 (a frequent co-author of Solymos) propose a pretty neat diagnostic tool to assess
estimability that is the by product of tricking a bayesian MCMC set up into Maximum Likelihood estimation for hierarchical models (see Lele et al 2007, 2010 and others. The keyword is ``Data Cloning"). You are one step away from using Solymos and Lele's ``Data Cloning" (DC) approach to get the full ML estimation working (See Lele et al 2010 and other Data Cloning papers). Now, this is relevant because using DC you can provide clear estimability diagnostics for the parameters of interest.  Are there parameters that are technically not estimable?  Given that you don't have informative priors, you are ``driving in the dark" so to speak if you are not certain if some of your parameters are un-identifiable (see Lele et al 2010 and discussion in Lele and Dennis 2009).  Note that I am NOT asking you to re-do all the analysis using a ML approach, I am just suggesting the fact that using DC to get the ML estimates one can, as a very useful by product, get very neat diagnostics regarding
the estimability of your parameters.  Again, see Lele et al 2010.  Implementing DC would only imply a simple modification to the programs you already have. And by the way, Solymos has a DC package easy to use.
\begin{quote}
\adam{
\begin{itemize}
\item The Solymos package is built for BUGS and JAGS.  Most of the tools are easily hand-coded, I think.
\item To simplify our lives, we could choose to attempt this for only quick-running models on the actual data... there's no reason to suspect identifiability issues for the gamma TTDD that would not appear for the Weibull or lognormal.
\item Technical note: they suggest using narrower priors as the number of clones increases.  Likewise, we could input posterior means as initial values.
\end{itemize}
}
\jarad{I'm not inclined to do this.}
\end{quote}

@@

One thing that the authors could do is to take the reader by the hand by building equation 1 bit by bit: present first the simplest version of equation 1 and what is customary to date, and then add modifications/hierarchies plus text and build the model with explicit equations and little by little arrive to the final product, which is the current equation 1. Nothing better than clear simple math aided by short explanations to present a model unambiguously.
\begin{quote}
\adam{
This potentially requires a lot of rewiring.  Two options are: (i) highlight each component of Equation (1) as we address it in the text, because we do address them all, or (ii) at the end of Section 3.2, provide a simpler version of the model without mixtures and without covariates.
}
\jarad{But it appears the reviewer wants it up front. 
I need to go back and reread to determine if this seems feasible. 
I thought we did start without mixtures and without covariates.}
\end{quote}

@@

The authors could have crafted a beautiful and very telling graph (Figure 1) representing the multinomial intervals, the TTDD, the right censoring due to the end of the observation period, and the increased detection probability mixture component. Such figure (which is what I did by hand as I was reading the manuscript) would be a very useful guide in the reading of the paper.
\begin{quote}
We have crafted a beautiful and very telling graph (Figure 1).
\jarad{Stay away from sass.}
\end{quote}

@@

\jarad{Ran out of steam?}

@@

\jarad{I missed this previously.}
